The Sloan Digital Sky Survey Two-dimensional Spectroscopic Reduction Pipeline

### 2012-06-19 doc/paper/paper2d.tex
Update.

\title{The Sloan Digital Sky Survey: Two-dimensional Spectroscopic Reduction Pipeline}

\author{
Scott Burles\altaffilmark{\ref{email}}
and David J. Schlegel\altaffilmark{\ref{Princeton}}
}
\altaffiltext{1}{sburles@gmail.com \label{sburles}}
\altaffiltext{2}{Princeton University Observatory, Peyton Hall, Princeton,
NJ 08544 \label{Princeton}}

@@

\begin{abstract}
The Sloan Digital Sky Survey (SDSS) fiber-fed spectrographs have been used
to obtain spectra for ~$600,000$ galaxies, quasars and stars
with a resolution of ~2000 over the wavelength range 3800\AA -9600\AA.
We present the pipeline design and algorithms for reducing the raw data to
wavelength-calibrated, flux-calibrated one-dimensional spectra.
It is designed to automatically reduce survey quality data sets and output 
uniformly processed results.  No human intervention is required in any step 
of the pipeline beyond invoking the top level command pointing to the 
applicable data.
The wavelength calibration is shown to have an accuracy of $3 \kms$,
and the broad-band spectrophotometry to have an accuracy of better than $6\%$.
What about reported errors and their accuracy/robustness?
The reported errors capture the statistical and the expected have

@@

%------------------------------------------------------------------------------
% INTRODUCTION
%------------------------------------------------------------------------------
\section{INTRODUCTION}
\label{sec_intro}
The Sloan Digital Sky Survey (SDSS) is a 5-band photometric survey of
approximately $10,000$ square degrees of the Northern sky and a concurrent
redshift survey of up to a million galaxies and 100,000 quasars
selected from the imaging survey (York et al 2000).  The primary 
science goals of the project are to provide the data to investigate the
large scale structure of the Universe as well as other extragalactic and 
galactic science.

@@

\begin{table}
\begin{tabular}{ccc}
\hline
DR &  2d version & rerun ? \\
\hline
\hline
DR2 & 4\_10\_4  & 23 \\
DR3 & 4\_10\_4  & 23 \\
DR4 & 4\_10\_4  & 23 \\
DR5 & 4\_10\_4 & 23 \\
DR6 & 5\_3\_3  & 25 \\
DR7 & 5\_3\_12 & 26 \\
DR8 & 5\_3\_12 & 26 \\
\hline
\end{tabular}
\end{table}

>>>>

\title{The Sloan Digital Sky Survey Two-dimensional Spectroscopic Reduction Pipeline}

\author{
David J. Schlegel\altaffilmark{\ref{LBL}}
and Scott Burles\altaffilmark{\ref{email}}
}
\altaffiltext{1}{sburles@gmail.com \label{sburles}}
\altaffiltext{2}{Lawrence Berkeley National Laboratory,
1 Cycltotron Road, Berkeley, CA 94720 \label{LBL}}

@@

\begin{abstract}
We describe the data reduction pipeline for the spectroscopic data
reduction for the optical spectrographs on the Sloan Telescope.
These spectrographs were used to obtain spectra of 900,000 galaxies,
100,000 QSOs and 300,000 stars during the Sloan Digital Sky Survey
(SDSS) and Sloan Digital Sky Survey II (SDSS-II) from 1999 through 2009.
The instrument was rebuilt in 2009 for the Sloan Digital Sky Survey III
(SDSS-III) and will be used to complete a survey of an additional
1.35 million galaxies and 200,000 QSOs by 2014.
We present the pipeline design and algorithms for reducing the raw
data to wavelength-calibrated, flux-calibrated one-dimensional spectra
and per-pixel error estimates.
The pipeline has been used to automatically reduce the survey data
sets and output uniformly-processed results.
The wavelength calibration is demonstrated to have an accuracy
of $3\kms$ and the broad-band spectrophotometry to have an accuracy
of $6\%$ for SDSS-I and II and $9\%$ for SDSS-III.
These spectra are of sufficient quality to allow automated redshift
fits to $99\%$ of the galaxies in SDSS-I and II and $\98\%$ of the much
fainter galaxies in SDSS-III, and catestrophic redshift errors
of much less than $1\%$ in all cases.

@@

%------------------------------------------------------------------------------
% INTRODUCTION
%------------------------------------------------------------------------------
\section{INTRODUCTION}
\label{sec_intro}
The Sloan Telescope was designed to provide 5-band imaging
of approximately one-quarter of the sky and spectroscopic follow-up
of millions of galaxies, QSOs and stars in that same footprint.
The original SDSS operated from 1998-2005 and
imaged 7600 deg$^2$ with spectroscopic follow-up of ???.
SDSS-II operated from 2005-2008 as a combination of three programs:
spectroscopic follow-up in the remainder of the original SDSS footprint,
a supernova search from repeated imaging of an equatorial stripe,
and imaging and spectroscopic follow-up for Galactic stars (SEGUE).

The SDSS spectroscopic pipeline (idlspec2d) was written in support
of the original galaxy and QSO redshift survey.
This pipeline was modified to provide more accurate redshifts
and spectrophotometry in support of the SEGUE program in SDSS-II.
The pipeline was again modified to support the rebuilt spectrographs
for the Baryon Oscillation Spectroscopic Survey (BOSS) in SDSS-III,
which is observing targets 2.5 mag fainter than SDSS.
Finally, many of the algorithms and code have been used for
the spectroscopic reductions of the DEEP2 survey.

The pipeline is written in the IDL language with some C bindings.
It makes use of the Goddard IDL astronomy library.
The code and all previous versions are publicly-available
through the SDSS-III project at {sdss3.org}.

The goal of the SDSS spectroscopic pipeline is to produce the
best estimate of spectral intensity as a function of wavelength.

@@

\begin{table}
\begin{tabular}{ccc}
\hline
Data Release & Instrument & Code version ? \\
\hline
\hline
DR1 & SDSS & ?   \\
DR2 & SDSS & v4\_10\_4   \\
DR3 & SDSS & v4\_10\_4   \\
DR4 & SDSS & v4\_10\_4   \\
DR5 & SDSS & v4\_10\_4  \\
DR6 & SDSS & v5\_3\_3   \\
DR7 & SDSS & v5\_3\_12  \\
DR8 & SDSS & v5\_3\_12  \\
DR9 & BOSS & ? \\
\hline
\end{tabular}
\end{table}
